---
title: "BellaBeat Analysis"
author: "Juan"
date: "2023-03-20"
output: html_document
---

#1. Install and load libraries
```{r}
install.packages("tidyverse")
install.packages("readr")
install.packages("dplyr")
install.packages("lubridate")
install.packages("hms")
install.packages("janitor")
install.packages("ggplot2")
install.packages("ggpubr")
install.packages("cowplot")
install.packages("scales")
library(tidyverse)
library(readr)
library(dplyr)
library(lubridate)
library(hms)
library(janitor)
library(ggplot2)
library(ggpubr)
library(cowplot)
library(scales)
```

#2. Upload datasets 
```{r}
daily_activity <- read.csv("dailyActivity_merged.csv")
daily_sleep <- read.csv("sleepDay_merged.csv")
hourly_calories <- read.csv("hourlyCalories_merged.csv")
hourly_steps <- read.csv("hourlySteps_merged.csv")
hourly_intensities <- read.csv("hourlyIntensities_merged.csv")
heart_rate_sec <- read.csv("heartrate_seconds_merged.csv")
```


#3. Transform column names to a consistent and clean format
```{r}
daily_activity <- clean_names(daily_activity)
daily_sleep <- clean_names(daily_sleep)
heart_rate_sec <- clean_names(heart_rate_sec)  
hourly_calories <- clean_names(hourly_calories)
hourly_steps <- clean_names(hourly_steps)
hourly_intensities <- clean_names(hourly_intensities)
```

  
#4. Ensure that id length is consistent 
```{r}
id_length1<- daily_activity %>%
  group_by(id) %>% 
  reframe(id_len = nchar(id)) %>% 
  distinct(id, id_len) 
id_length1

id_length2<- daily_sleep %>%
  group_by(id) %>% 
  reframe(id_len = nchar(id)) %>% 
  distinct(id, id_len)  
id_length2

id_length3<- hourly_calories %>%
  group_by(id) %>% 
  reframe(id_len = nchar(id)) %>% 
  distinct(id, id_len)  
id_length3

id_length4<- hourly_steps %>%
  group_by(id) %>% 
  reframe(id_len = nchar(id)) %>% 
  distinct(id, id_len)  
id_length4

id_length5<- hourly_intensities %>%
  group_by(id) %>% 
  reframe(id_len = nchar(id)) %>% 
  distinct(id, id_len) 
id_length5

id_length6<- heart_rate_sec %>%
  group_by(id) %>% 
  reframe(id_len = nchar(id)) %>% 
  distinct(id, id_len)
id_length6 
```

#5.1 Change "date" column format to <date> format
```{r}
daily_activity <- daily_activity %>% 
  mutate(activity_date = mdy(activity_date)) %>% 
  mutate(activity_date = as.Date(activity_date)) %>% 
  rename(date = activity_date)

daily_sleep <- daily_sleep %>% 
  mutate(sleep_day = mdy_hms(sleep_day)) %>% 
  mutate(sleep_day = as.Date(sleep_day)) %>% 
  rename(date = sleep_day)
```

#5.2 Transform "date" column into "date" and "time" columns, each with appropiate format
```{r}
hourly_calories <- hourly_calories %>% 
  mutate(activity_hour = mdy_hms(activity_hour)) %>% 
  separate(activity_hour,into=c("date", "time"), sep=" ") %>% 
  mutate(date=as.Date(date), time=as_hms(time))

hourly_steps <- hourly_steps %>% 
  mutate(activity_hour = mdy_hms(activity_hour)) %>% 
  separate(activity_hour,into=c("date", "time"), sep=" ") %>% 
  mutate(date=as.Date(date), time=as_hms(time))

hourly_intensities <- hourly_intensities %>% 
  mutate(activity_hour = mdy_hms(activity_hour)) %>% 
  separate(activity_hour,into=c("date", "time"), sep=" ") %>% 
  mutate(date=as.Date(date), time=as_hms(time))

heart_rate_sec <- heart_rate_sec %>% 
  mutate(time = mdy_hms(time)) %>% 
  separate(time,into=c("date", "time"), sep=" ") %>% 
  mutate(date=as.Date(date), time=as_hms(time))
```

#6.1 Check datasets with a limited number of columns for duplicates
```{r}
sum(duplicated(heart_rate_sec))

sum(duplicated(hourly_calories))

sum(duplicated(hourly_intensities))

sum(duplicated(hourly_steps))
```

#6.2 Check datasets with multiple columns for duplicates
```{r}
dup_entries_count1 <- daily_activity %>%
  mutate(combined = paste(id, as.character(date), sep = "_")) %>%
  filter(duplicated(combined)) %>% 
  nrow()
dup_entries_count1
 
dup_entries_count2 <- daily_sleep %>%
  mutate(combined = paste(id, as.character(date), sep = "_")) %>%
  filter(duplicated(combined)) %>% 
  nrow()
dup_entries_count2
```

#6.3 Remove duplicates in the daily_sleep dataset
```{r}
daily_sleep <- daily_sleep %>%
  distinct() 

dup_entries_count3 <- daily_sleep %>%
  mutate(combined = paste(id, as.character(date), sep = "_")) %>%
  filter(duplicated(combined)) %>% 
  nrow()
dup_entries_count3
```

#7. Find missing and N.A. values
```{r}
missing_values1 <- daily_activity %>%
  mutate_if(is.Date, as.character) %>% 
  summarize_all(~sum(is.na(.) | . == ""))
missing_values1 

missing_values2 <- daily_sleep %>%
  mutate_if(is.Date, as.character) %>% 
  summarize_all(~sum(is.na(.) | . == ""))
missing_values2

missing_values3 <- hourly_calories %>%
  mutate_if(is.Date, as.character) %>% 
  summarize_all(~sum(is.na(.) | . == ""))
missing_values3

missing_values4 <- hourly_steps %>%
  mutate_if(is.Date, as.character) %>% 
  summarize_all(~sum(is.na(.) | . == ""))
missing_values4

missing_values5 <- hourly_intensities %>%
  mutate_if(is.Date, as.character) %>% 
  summarize_all(~sum(is.na(.) | . == ""))
missing_values5

missing_values6 <- heart_rate_sec %>%
  mutate_if(is.Date, as.character) %>% 
  summarize_all(~sum(is.na(.) | . == ""))
missing_values6
```

#8.1 Categorizing User Activity Level in a dataframe
```{r}
user_type <- daily_activity %>% 
  group_by(id) %>% 
  summarise(avg_steps=mean(total_steps), sd_steps=sd(total_steps), lower_sd_steps=mean(total_steps)-sd(total_steps)) %>% 
  mutate(user_type = case_when(
    avg_steps < 5000 ~ "sedentary",
    avg_steps >= 5000 & avg_steps < 7500 ~ "lightly active", 
    avg_steps >= 7500 & avg_steps < 10000 ~ "fairly active", 
    avg_steps >= 10000 ~ "very active"
  )) %>% 
  arrange(avg_steps)
head(user_type)
```

#8.2 Plot Activity Levels in a Pie Graph
```{r}
activity_levels_percentage <- user_type %>%
  group_by(user_type) %>%
  summarize(count = n()) %>%
  mutate(percentage = percent(count / sum(count)))
head(activity_levels_percentage)

ggplot(activity_levels_percentage, mapping=aes(x="", y=percentage, fill=user_type)) +
  geom_bar(stat="identity") +
  coord_polar(theta = "y") +
  theme_minimal() +
  geom_text(aes(label=paste(user_type, "\n", percentage)),
            position = position_stack(vjust=0.5),
            show.legend = none, fontface="bold", size=3) +
  theme(axis.title.x= element_blank(),
        axis.title.y = element_blank(),
        panel.border = element_blank(), 
        panel.grid = element_blank(), 
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        plot.title = element_text(hjust = 0.5, size=14, face = "bold"))+
  guides(fill = FALSE)+
  labs(title = "User Activity Level Category")
```

#9 Join tables for in-depth analysis 
```{r}
daily_data <- left_join(daily_activity, daily_sleep, by = c("id", "date")) %>% 
  drop_na()

hourly_data <- hourly_calories %>% 
  left_join(hourly_steps, by = c("id", "date", "time")) %>% 
  left_join(hourly_intensities, by = c("id", "date", "time")) %>% 
  drop_na()
```

#10.1 Users usage of smart device
```{r}
total_days <- daily_data %>% 
  summarise(total_days = n_distinct(date))
total_days

user_usage <- daily_data %>%
  group_by(id) %>% 
  reframe(days_used = n_distinct(date), hours_used=round((very_active_minutes+
              fairly_active_minutes+lightly_active_minutes+sedentary_minutes)/60,1)) %>% 
  mutate(days_usage_percentage=round(days_used/total_days$total_days*100), 
        hours_usage_percentage=round(hours_used/24*100)) %>% 
  distinct(id, .keep_all = TRUE)
head(user_usage)
```

#10.2 Correlation between days and hours of usage
```{r}
ggplot(user_usage, aes(x=days_usage_percentage, y=hours_usage_percentage)) +
  geom_jitter() +
  geom_smooth(color = "red") + 
  labs(title = "Days vs Hours Usage", x = "Days Usage", y= "Hours Usage") +
  theme(plot.title = element_text(hjust=0.5))
```

#11.1 Users Daily activity levels
```{r}
weekly_active <- daily_data %>% 
  mutate(day_of_week = wday(date, label= TRUE)) %>% 
  group_by(day_of_week) %>% 
  summarise(avg_daily_steps = mean(total_steps))
head(weekly_active)
```

#11.2 Users Daily activity levels Plot
```{r}
daily_steps_plot <- ggplot(weekly_active, mapping=aes(x=day_of_week, y=avg_daily_steps))+ 
  geom_bar(stat="identity", fill="#F8766D")+
  geom_hline(yintercept = mean(weekly_active$avg_daily_steps), color = "black", size=1.2) +
  annotate("text", x=.95, y=mean(weekly_active$avg_daily_steps), 
           label=paste0("Avg:", round(mean(weekly_active$avg_daily_steps))), size=3.3, vjust=-0.5) +
  labs(title="User Steps by Day of Week", x="Day of Week", y= "Average User Steps")+
  theme(plot.title = element_text(hjust=0.5))
print(daily_steps_plot)
```

#12.1 Users Hourly activity
```{r}
daily_active <- hourly_data %>% 
  group_by(time) %>% 
  summarise(avg_hourly_steps = mean(step_total))
head(daily_active)
```

#12.2 Users Hourly activity Plot
```{r}
hourly_steps_plot <- ggplot(daily_active, mapping=aes(x=time, y=avg_hourly_steps)) +
  geom_bar(stat="identity", fill="#619CFF")+
  geom_hline(yintercept = mean(daily_active$avg_hourly_steps), color = "black", size=1.2) +
  annotate("text", x=1, y=mean(daily_active$avg_hourly_steps),
           label=paste0("Avg:", round(mean(daily_active$avg_hourly_steps))), size=3.3, vjust=-0.5, hjust=0.4) +
  labs(title="", x="", y="") +
  theme(plot.title = element_text(size=7,hjust=0.5), axis.text.x = element_text(angle=45, hjust=1))
print(hourly_steps_plot)
```

#13.1 Daily Step Comparison Dataframes - Very Active vs Less Active
```{r}
very_active_daily <- daily_data %>%
  group_by(id) %>% 
  summarise(avg_daily_steps = mean(total_steps)) %>%
  filter(avg_daily_steps > 10000) %>% 
  select(id) %>%
  inner_join(daily_data, by = "id") %>% 
  mutate(day_of_week = wday(date, label= TRUE)) %>% 
  group_by(day_of_week) %>% 
  summarise(avg_daily_steps = mean(total_steps))
View(very_active_daily)

less_active_daily <- daily_data %>%
  group_by(id) %>% 
  summarise(avg_daily_steps = mean(total_steps)) %>%
  filter(avg_daily_steps < 10000) %>% 
  select(id) %>%
  inner_join(daily_data, by = "id") %>% 
  mutate(day_of_week = wday(date, label= TRUE)) %>% 
  group_by(day_of_week) %>% 
  summarise(avg_daily_steps = mean(total_steps))
View(less_active_daily)
```

#13.2 Daily Step Comparison Plot - Very Active vs Less Active
```{r}
very_active_daily_plot <- ggplot(very_active_daily, mapping=aes(x=day_of_week, y=avg_daily_steps))+ 
  geom_bar(stat="identity", fill="#F8766D")+
  geom_hline(yintercept = mean(very_active_daily$avg_daily_steps), color = "black", size=1) +
  annotate("text", x=.95, y=mean(very_active_daily$avg_daily_steps), 
           label=paste0("Avg:", round(mean(very_active_daily$avg_daily_steps))), size=2.5, vjust=-0.5, hjust=-1) +
  labs(title="Very Active Daily Average Steps", x="Day of Week", y= "Very Active User Steps")+
  theme(plot.title = element_text(hjust=0.5, size=6), axis.text.x = element_text(angle=45, hjust=1), 
        axis.title = element_text(size=8))+
  ylim(0, 15000)

less_active_daily_plot <- ggplot(less_active_daily, mapping=aes(x=day_of_week, y=avg_daily_steps))+ 
  geom_bar(stat="identity", fill="#F8766D")+
  geom_hline(yintercept = mean(less_active_daily$avg_daily_steps), color = "black", size=1) +
  annotate("text", x=.95, y=mean(less_active_daily$avg_daily_steps), 
           label=paste0("Avg:", round(mean(less_active_daily$avg_daily_steps))), size=2.5, vjust=-0.5, hjust=-1) +
  labs(title="Less Active Daily Average Steps", x="Day of Week", y= "Less Active User Steps")+
  theme(plot.title = element_text(hjust=0.5, size=6), axis.text.x = element_text(angle=45, hjust=1), 
        axis.title = element_text(size=8)) + 
  ylim(0, 15000)

combined_daily_plots <- ggarrange(very_active_daily_plot, less_active_daily_plot, ncol = 2) %>% 
  ggdraw() +
  draw_label("") 
print(combined_daily_plots)
```

#14.1 Hourly Step Comparison Dataframes - Very Active vs Less Active
```{r}
very_active_hourly <- daily_data %>%
  group_by(id) %>% 
  summarise(avg_daily_steps = mean(total_steps)) %>%
  filter(avg_daily_steps > 10000) %>% 
  select(id) %>%
  inner_join(hourly_data, by = "id") %>% 
  group_by(time) %>% 
  summarise(avg_hourly_steps = mean(step_total))
View(very_active_hourly)

less_active_hourly <- daily_data %>%
  group_by(id) %>% 
  summarise(avg_daily_steps = mean(total_steps)) %>%
  filter(avg_daily_steps < 10000) %>% 
  select(id) %>%
  inner_join(hourly_data, by = "id") %>% 
  group_by(time) %>% 
  summarise(avg_hourly_steps = mean(step_total))
View(less_active_hourly)
```

#14.2 Hourly Step Comparison Plot - Very Active vs Less Active
```{r}
very_active_hourly_plot <- ggplot(very_active_hourly, mapping=aes(x=time, y=avg_hourly_steps)) +
  geom_bar(stat="identity", fill="#619CFF")+
  geom_hline(yintercept = mean(very_active_hourly$avg_hourly_steps), color = "black", size=1) +
  annotate("text", x=1, y=mean(very_active_hourly$avg_hourly_steps),
           label=paste0("Avg:", round(mean(very_active_hourly$avg_hourly_steps))), size=2.5, vjust=-0.5, hjust=0) +
  labs(title="Very Active Hourly Average Steps", x="Time", y= "Very Active User Steps")+
    theme(plot.title = element_text(hjust=0.5, size=6), axis.text.x = element_text(angle=45, hjust=1), 
        axis.title = element_text(size=8)) +
  scale_y_continuous(limits = c(0, 1400), breaks = seq(0, 1400, by = 200))
print(very_active_hourly_plot)

less_active_hourly_plot <- ggplot(less_active_hourly, mapping=aes(x=time, y=avg_hourly_steps)) +
  geom_bar(stat="identity", fill="#619CFF")+
  geom_hline(yintercept = mean(less_active_hourly$avg_hourly_steps), color = "black", size=1) +
  annotate("text", x=1, y=mean(less_active_hourly$avg_hourly_steps),
           label=paste0("Avg:", round(mean(less_active_hourly$avg_hourly_steps))), size=2.5, vjust=-0.5, hjust=0.2) +
  labs(title="Less Active Hourly Average Steps", x="Time", y= "Less Active User Steps")+
  theme(plot.title = element_text(hjust=0.5, size=6), axis.text.x = element_text(angle=45, hjust=1), 
        axis.title = element_text(size=8))+
  scale_y_continuous(limits = c(0, 1400), breaks = seq(0, 1400, by = 200))
print(less_active_hourly_plot)

combined_hourly_plots <- ggarrange(very_active_hourly_plot, less_active_hourly_plot, ncol = 2) %>% 
  ggdraw() +
  draw_label("") 
print(combined_hourly_plots)
```

#15.1 Users Average Sleep Duration Dataframe
```{r}
asleep_data <- daily_data %>% 
  mutate(day_of_week = wday(date, label= TRUE)) %>% 
  group_by(day_of_week) %>% 
  summarise(avg_sleep_hr = mean(total_minutes_asleep)/60)
head(asleep_data)
```

#15.2 Users Average Sleep Duration Plot
```{r}
ggplot(asleep_data, mapping=aes(x=day_of_week, y=avg_sleep_hr))+
  geom_bar(stat="identity", fill="palegreen4")+
  geom_hline(yintercept = mean(asleep_data$avg_sleep_hr), color = "black", size=1) +
  annotate("text", x=1, y=mean(asleep_data$avg_sleep_hr),
           label=paste0("Avg:", round(mean(asleep_data$avg_sleep_hr))), size=2.5, vjust=-0.5, hjust=-1) +
  labs(title="Daily Hours of Sleep", x="Day of the Week", y="Asleep Hours")+
  scale_y_continuous(breaks = seq(0, 8, by = 1))+ 
  theme(plot.title = element_text(size=10, hjust=0.5), axis.text.x = element_text(hjust=1))
```

#16.1 Users Average Fall Asleep Duration Dataframe
```{r}
falling_asleep_data <- daily_data %>% 
  mutate(day_of_week = wday(date, label= TRUE)) %>% 
  group_by(day_of_week) %>% 
  summarise(avg_falling_sleep = (mean(total_time_in_bed)-mean(total_minutes_asleep))/60)
head(falling_asleep_data)
```

#16.2 Users Average Fall Asleep Duration Plot
```{r}
ggplot(falling_asleep_data, mapping=aes(x=day_of_week, y=avg_falling_sleep)) +
  geom_bar(stat="identity", fill="paleturquoise4") +
  geom_hline(yintercept = mean(falling_asleep_data$avg_falling_sleep), color = "black", size=1) +
  annotate("text", x=1, y=mean(falling_asleep_data$avg_falling_sleep),
           label=paste0("Avg:", round(mean(falling_asleep_data$avg_falling_sleep),2), " hr"), size=3, vjust=-0.5, hjust=-1) +
  labs(title="Daily Hours To Fall Asleep", x="Day of the Week", y="Hours To Fall Asleep") +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1)) + 
  theme(plot.title = element_text(size=10, hjust=0.5), axis.text.x = element_text(hjust=1))
```

#17.1 Very Active & Less Active sleep and Fall Asleep Dataframes
```{r}
very_active_sleep <- daily_data %>%
  group_by(id) %>% 
  summarise(avg_daily_steps = mean(total_steps)) %>%
  filter(avg_daily_steps > 10000) %>% 
  select(id) %>%
  inner_join(daily_data, by = "id") %>% 
  mutate(day_of_week = wday(date, label= TRUE)) %>% 
  group_by(day_of_week) %>% 
  summarise(avg_sleep_hr = mean(total_minutes_asleep)/60, 
            avg_falling_asleep = (mean(total_time_in_bed)-mean(total_minutes_asleep))/60)
View(very_active_sleep)

less_active_sleep <- daily_data %>%
  group_by(id) %>% 
  summarise(avg_daily_steps = mean(total_steps)) %>%
  filter(avg_daily_steps < 10000) %>% 
  select(id) %>%
  inner_join(daily_data, by = "id") %>% 
  mutate(day_of_week = wday(date, label= TRUE)) %>% 
  group_by(day_of_week) %>% 
  summarise(avg_sleep_hr = mean(total_minutes_asleep)/60, 
            avg_falling_asleep = (mean(total_time_in_bed)-mean(total_minutes_asleep))/60)
View(less_active_sleep)
```

#17.2 Very Active vs Less Active Sleep Comparison Plot
```{r}
combined_sleep_plots <- ggarrange(
  ggplot(very_active_sleep, mapping=aes(x=day_of_week, y=avg_sleep_hr))+
    geom_bar(stat="identity", fill="palegreen4")+
    geom_hline(yintercept = mean(very_active_sleep$avg_sleep_hr), color = "black", size=1) +
    annotate("text", x=1, y=mean(very_active_sleep$avg_sleep_hr),
             label=paste0("Avg:", round(mean(very_active_sleep$avg_sleep_hr),2)), size=2.5, vjust=-0.5, hjust=-0.8) +
    labs(title="Very Active Sleep Hours", x="Day of the Week", y="Asleep Hours")+
    scale_y_continuous(breaks = seq(0, 8, by = 1))+ 
    theme(plot.title = element_text(size=10, hjust=0.5), axis.text.x = element_text(hjust=1)),
  ggplot(less_active_sleep, mapping=aes(x=day_of_week, y=avg_sleep_hr))+
    geom_bar(stat="identity", fill="palegreen4")+
    geom_hline(yintercept = mean(less_active_sleep$avg_sleep_hr), color = "black", size=1) +
    annotate("text", x=1, y=mean(less_active_sleep$avg_sleep_hr),
             label=paste0("Avg:", round(mean(less_active_sleep$avg_sleep_hr),2)), size=2.5, vjust=-0.5, hjust=0.5) +
    labs(title="Less Active Sleep Hours", x="Day of the Week", y="Asleep Hours")+
    scale_y_continuous(breaks = seq(0, 8, by = 1))+ 
    theme(plot.title = element_text(size=10, hjust=0.5), axis.text.x = element_text(hjust=1)),
  nrow = 2)
print(combined_sleep_plots)
```

#17.3 Very Active vs Less Active Fall Asleep Comparison Plot
```{r}
combined_falling_sleep_plots <- ggarrange(
  ggplot(very_active_sleep, mapping=aes(x=day_of_week, y=avg_falling_asleep)) +
    geom_bar(stat="identity", fill="paleturquoise4") +
    geom_hline(yintercept = mean(very_active_sleep$avg_falling_asleep), color = "black", size=1) +
    annotate("text", x=1, y=mean(very_active_sleep$avg_falling_asleep),
             label=paste0("Avg:", round(mean(very_active_sleep$avg_falling_asleep),2), " hr"), size=3, vjust=-0.5, hjust=-0.2) +
    labs(title="Very Active Falling Asleep Hours", x="Day of the Week", y="Hours To Fall Asleep") +
    scale_y_continuous(breaks = seq(0, 1, by = 0.1)) + 
    theme(plot.title = element_text(size=10, hjust=0.5), axis.text.x = element_text(hjust=1)),
  ggplot(less_active_sleep, mapping=aes(x=day_of_week, y=avg_falling_asleep)) +
    geom_bar(stat="identity", fill="paleturquoise4") +
    geom_hline(yintercept = mean(less_active_sleep$avg_falling_asleep), color = "black", size=1) +
    annotate("text", x=1, y=mean(less_active_sleep$avg_falling_asleep),
             label=paste0("Avg:", round(mean(less_active_sleep$avg_falling_asleep),2), " hr"), size=3, vjust=-0.5, hjust=-1.2) +
    labs(title="Less Active Falling Asleep Hours", x="Day of the Week", y="Hours To Fall Asleep") +
    scale_y_continuous(breaks = seq(0, 1, by = 0.1)) + 
    theme(plot.title = element_text(size=10, hjust=0.5), axis.text.x = element_text(hjust=1)),
  nrow = 2)
print(combined_falling_sleep_plots)
```

#18 Heart Rate Analysis
```{r}
very_active_heart_rate <- daily_data %>%
  group_by(id) %>% 
  summarise(avg_daily_steps = mean(total_steps)) %>%
  filter(avg_daily_steps > 10000) %>% 
  select(id) %>%
  inner_join(heart_rate_sec, by = "id") %>%
  distinct(id)
```

#19.1 Very Active and Less Active Evening Steps Dataframes
```{r}
very_active_evening_steps <- daily_data %>%
  group_by(id) %>% 
  summarise(avg_daily_steps = mean(total_steps)) %>%
  filter(avg_daily_steps > 10000) %>% 
  select(id) %>%
  inner_join(hourly_data, by = "id") %>% 
  filter(between(hour(time), 18, 24)) %>% 
  group_by(time) %>% 
  summarise(avg_hourly_steps = mean(step_total))
View(very_active_evening_steps)

less_active_evening_steps <- daily_data %>%
  group_by(id) %>% 
  summarise(avg_daily_steps = mean(total_steps)) %>%
  filter(avg_daily_steps < 10000) %>% 
  select(id) %>%
  inner_join(hourly_data, by = "id") %>% 
  filter(between(hour(time), 18, 24)) %>% 
  group_by(time) %>% 
  summarise(avg_hourly_steps = mean(step_total))
View(less_active_evening_steps)
```

#19.2 Very Active and Less Active Evening Steps Line Plot
```{r}
ggplot() +
  geom_line(data = very_active_evening_steps, aes(x = time, y = avg_hourly_steps, color = "Very Active")) +
  geom_line(data = less_active_evening_steps, aes(x = time, y = avg_hourly_steps, color = "Less Active")) +
  geom_point(data = very_active_evening_steps, aes(x = time, y = avg_hourly_steps), size = 2.5, shape = 21, fill = "#FFFFFF", color = "red") +
  geom_point(data = less_active_evening_steps, aes(x = time, y = avg_hourly_steps), size = 2.5, shape = 21, fill = "#FFFFFF", color = "blue") +
  labs(title = "Evening Steps", x = "Time", y = "Average Steps") +
  scale_color_manual(values = c("Very Active" = "red", "Less Active" = "blue")) +
  scale_y_continuous(breaks = seq(50, 1400, by = 150)) +
  theme(plot.title = element_text(size = 10, hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
```

#20.1 Falling Asleep Vs Sleep Duration Correlation
```{r}
sleep_corr1 <- daily_data %>% 
  mutate(mins_to_fall_asleep = total_time_in_bed - total_minutes_asleep) %>%
  ggplot(mapping=aes(x=mins_to_fall_asleep/60, y=total_minutes_asleep/60)) +
  geom_jitter() +
  geom_smooth(color = "red") + 
  labs(title = "Hours Asleep vs Hours to Fall Asleep", x = "Hours to Fall Asleep", y= "Hours Asleep") +
  scale_y_continuous(breaks = seq(0, 14, by = 2)) + 
  scale_x_continuous(breaks = seq(0, 7, by = 1)) +
  theme(plot.title = element_text(hjust=0.5, size=14))
print(sleep_corr1)
```

#20.2 Falling Asleep Vs Sleep Duration Correlation (2 hours Time frame)
```{r}
sleep_corr2 <- daily_data %>% 
  mutate(mins_to_fall_asleep = total_time_in_bed - total_minutes_asleep) %>%
  filter(mins_to_fall_asleep < 60*2) %>% 
  ggplot(mapping=aes(x=mins_to_fall_asleep/60, y=total_minutes_asleep/60)) +
  geom_jitter() +
  geom_smooth(color = "red") + 
  labs(title = "Hours Asleep vs Hours to Fall Asleep", x = "Hours to Fall Asleep", y= "Hours Asleep") +
  scale_y_continuous(breaks = seq(0, 14, by = 2)) + 
  scale_x_continuous(breaks = seq(0, 7, by = 1)) +
  theme(plot.title = element_text(hjust=0.5, size=14))
print(sleep_corr2)
```

#21 Daily Steps vs Falling Asleep Correlation
```{r}
sleep_corr3 <- daily_data %>% 
  mutate(mins_to_fall_asleep = total_time_in_bed - total_minutes_asleep) %>%
  ggplot(mapping=aes(x=total_steps, y=mins_to_fall_asleep/60)) +
  geom_jitter() +
  geom_smooth(color = "red") + 
  labs(title = "Daily Steps vs Duration to Fall Asleep", x = "Daily Steps", y= "Duration to Fall Asleep") +
  scale_y_continuous(breaks = seq(0, 14, by = 2)) + 
  theme(plot.title = element_text(hjust=0.5, size=14))
print(sleep_corr3)

#22 Daily Steps Vs Asleep Duration Correlation
sleep_corr4 <- daily_data %>% 
  ggplot(mapping=aes(x=total_steps, y=total_minutes_asleep/60)) +
  geom_jitter() +
  geom_smooth(color = "red") + 
  labs(title = "Sleep Hours vs Total Steps", x = "Daily Steps", y= "Daily Asleep Hours") +
  scale_y_continuous(breaks = seq(0, 16, by = 2)) + 
  theme(plot.title = element_text(hjust=0.5, size=14),axis.text.x = element_text(hjust=0.5))
print(sleep_corr4)
```

#23 Intensity Activity Vs Asleep Duration Correlation
```{r}
user_avg_sleep <- daily_data %>% 
  group_by(id) %>% 
  summarise(avg_daily_asleep = mean(total_minutes_asleep/60))

user_avg_intensity <- hourly_data %>% 
  group_by(id) %>% 
  summarise(avg_daily_intensity = mean(total_intensity))

sleep_corr5 <- user_avg_sleep %>% 
  left_join(user_avg_intensity, by = c("id")) %>% 
  ggplot(mapping=aes(x=avg_daily_intensity, y=avg_daily_asleep)) +   
  geom_jitter() +
  geom_smooth(color = "red") + 
  labs(title = "Intensity vs Sleep", x = "Daily Average Itensity", y= "Daily Average Sleep") +
  theme(plot.title = element_text(hjust=0.5, size=14))
print(sleep_corr5)
```

#24 Burned Calories Vs Asleep Duration Correlation
```{r}
sleep_corr6 <- daily_data %>% 
  ggplot(mapping=aes(x=calories, y=total_minutes_asleep/60)) +
  geom_jitter() +
  geom_smooth(color = "red") + 
  labs(title = "Calories Burned vs Sleep Hours", x = "Calories Burned", y= "Daily Asleep Hours") +
  scale_y_continuous(breaks = seq(0,16, by = 2)) + 
  theme(plot.title = element_text(hjust=0.5, size=14))
print(sleep_corr6)
```

#25 User Activity Categories Vs Asleep Duration Correlations
```{r}
combined_plots_activity <- ggarrange(
  ggplot(daily_data, mapping=aes(x=very_active_minutes/60, y=total_minutes_asleep/60))+   
    geom_jitter() +
    geom_smooth(color = "red") + 
    labs(title = "Very Active vs Sleep", x = "Very Active Hours", y= "Sleep Hours") +
    theme(plot.title=element_text(hjust=0.5, size=11, face="bold"), 
          axis.title.x=element_text(size=9), axis.title.y=element_text(size=9),
          axis.text.x=element_text(size=6), axis.text.y=element_text(size=6)),
  ggplot(daily_data, mapping=aes(x=fairly_active_minutes/60, y=total_minutes_asleep/60))+   
    geom_jitter() +
    geom_smooth(color = "red") + 
    labs(title = "Fairly Active vs Sleep", x = "Fairly Active Hours", y= "") +
    theme(plot.title = element_text(hjust=0.5, size=11, face="bold"), 
          axis.title.x=element_text(size=9), axis.title.y=element_text(size=9),
          axis.text.x=element_text(size=6), axis.text.y=element_text(size=6)),
  ggplot(daily_data, mapping=aes(x=lightly_active_minutes/60, y=total_minutes_asleep/60))+   
    geom_jitter() +
    geom_smooth(color = "red") + 
    labs(title = "Lightly Active vs Sleep", x = "Lightly Active Hours", y= "Sleep Hours") +
    theme(plot.title = element_text(hjust=0.5, size=11, face="bold"), 
          axis.title.x=element_text(size=9), axis.title.y=element_text(size=9),
          axis.text.x=element_text(size=6), axis.text.y=element_text(size=6)),
  ggplot(daily_data, mapping=aes(x=sedentary_minutes/60, y=total_minutes_asleep/60))+   
    geom_jitter() +
    geom_smooth(color = "red") + 
    labs(title = "Sedentary vs Sleep", x = "Sedentary Hours", y= "") +
    theme(plot.title = element_text(hjust=0.5, size=11, face="bold"), 
          axis.title.x=element_text(size=9), axis.title.y=element_text(size=9),
          axis.text.x=element_text(size=6), axis.text.y=element_text(size=6)),
  nrow = 2, ncol = 2)
print(combined_plots_activity)
```

#26 Daily Steps Vs Calories Burned Correlation
```{r}
cal_corr1 <- daily_data %>% 
  ggplot(mapping=aes(x=total_steps, y=calories)) +
  geom_jitter() +
  geom_smooth(color = "red") + 
  labs(title = "Daily Steps vs Calories", x = "Daily Steps", y= "Calories Burned")+
  scale_x_continuous(breaks = seq(0, 25000, by = 3000)) + 
  theme(plot.title = element_text(hjust=0.5, size=14),
        axis.text.x = element_text(angle=45, hjust=1))
print(cal_corr1)
```

#27 Intensity Vs Calories Burned Correlation
```{r}
cal_corr2 <- hourly_data %>% 
  ggplot(mapping=aes(x=average_intensity, y=calories)) +
  geom_jitter() +
  geom_smooth(color = "red") + 
  labs(title = "Intensity vs Calories", x = "Hourly Average Itensity", y= "Hourly Calories Burned") +
  theme(plot.title = element_text(hjust=0.5, size=14),
        axis.text.x = element_text(hjust=1))
print(cal_corr2)
```


